Stage-Splitting a Modal Language

### 2014-07-03 paper/intro.tex
fixed brackets
wrote a bit of intro

We often find ourselves in situations where the input to a computation arrives in two parts, at different points in time.

The natural response is to try and do some part of the computation early, and the rest of the computation once the second input arrives.

Since the computation runs in two parts, the total cost can be split as well, into $m+n$.

Additionally, we can run the second stage multiple times.  Cost is now $m+bn$.

@@

Likewise, we can easily recover a stage-splitter from a partial evaluator.  If $p$ is a valid partial evaluator, then we can define a stage splitter $s$ such that $s(f)=(f_1,f_2)$, where
\begin{align*}
[f_1](x) &= p (f,x) \\
[f_2](l,y) &= [l] (y)
\end{align*}
This implicitly requires that the languages in which $f_1$ and $f_2$ are expressed are strong enough to write a partial evaluator, but that is the case in this paper.  A stage-splitter defined this way leaves much to be desired.  Firstly, partial evaluation of $f$ may be too expensive for the context in which $f_1$ needs to run.  Additionally, the intermediate data structure created this way may be much larger than necessary, as it would contain all of the residual code.

>>>>

It is a common observation that some computations can be performed efficiently by using {\em staging}, wherein the execution of a program is split over two or more discrete points in time ({\em i.e.} stages).  This most often arises when we want to evaluate a multivariate function several times, with some inputs dynamic (varying across invocations) and some static (constant across invocations).  Examples include,
\begin{itemize}
\item Finding the $i$th largest element in a list, where the list is constant and $i$ varies.
\item Computing the shading of a surfacing in a 3D rendering, where some contributors to the final pixel color can be approximated as constant over $4\times4$ pixel blocks.  
\item Running a language interpreter on a single program for multiple inputs.
\end{itemize}

The goal is to run programs in a way that doesn't repeat work unless it depends on the varying input.  There are two separate problems at play here.  The first is {\em binding-time analysis}, the problem of determining {\em what} portions of the program to run statically versus dynamically.  After that is {\em staged execution}, the problem of {\em how} to execute in a staged way. We  consider only the latter in this paper.  That is, we assume that programs are given in a representation, covered later, that fully determines staged execution.

We consider two possible models of staged execution,

\TODO partial eval

\TODO stage-splitting

@@

Likewise, we can easily recover a stage-splitter from a partial evaluator.  If $p$ is a valid partial evaluator, then we can define a stage splitter $s$ such that $s(f)=(f_1,f_2)$, where
\begin{align*}
\llbracket f_1 \rrbracket (x) &= p (f,x) \\
\llbracket f_2 \rrbracket (l,y) &= [l] (y)
\end{align*}
This implicitly requires that the languages in which $f_1$ and $f_2$ are expressed are strong enough to write a partial evaluator, but that is the case in this paper.  A stage-splitter defined this way leaves much to be desired.  Firstly, partial evaluation of $f$ may be too expensive for the context in which $f_1$ needs to run.  Additionally, the intermediate data structure created this way may be much larger than necessary, as it would contain all of the residual code.
